---
title: "HostSwitch"
author: "Valeria Trivellona, Sabrina B.L. Araujo, Bernd Panassiti"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for the HostSwitch package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "The 'HostSwitch' package uses a simulation-based approach to investigate host switches by a parasite. The individual based model is based on the concepts of Stockholm paradigm and ecological fitting. The orginial model is published by Araujo et al. 2015. Understanding Host-Switching by Ecological Fitting [https://doi.org/10.1371/journal.pone.0139225](https://doi.org/10.1371/journal.pone.0139225). The package provides an R-friendly and modified version of this model which can be applied to different host-parasite scenarios. \n \n \n"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8.5, fig.height=5.5, fig.align='center', warning=FALSE, message=FALSE, cache = T)
```

```{r setup,echo = F, message = F}
library(HostSwitch)
```

# Motivation 

The vector-born pathogens are part of a complex interaction between vector-pathogen-host. This package focues on the 


# Using HostSwitch

## Installing, loading and citing the package

If you haven't installed the package yet, please install the recent release using GitHub

```{r, eval = F}
devtools::install_github(repo = "berndpanassiti/HostSwitch")
```

Loading and citation

```{r}
library(HostSwitch)
citation("HostSwitch")
```


## Run the model
Simulate host switches by parasites
```{r,eval=F}
HostSwitch_simulated_quantities = simHostSwitch(K=100,b=10, mig=0.001, sdm=0.2, sigma=1, pRes_min=1, pRes_max=10, n_generation=200)
HostSwitch_simulated_quantities
```

The function generates a list containing all quantities of intest which can be used for summary statistics or plots. Quantities of interests are **phenotpyes** of resource (original host, 'pRes_sim'), new resource (new host, 'pRes_new_sim') and of individual parasites ('pInd'). These parameters are available for each generation step.


## Plot the model
```{r,eval=F}
gg1 = plotHostSwitch(HostSwitch_simulated_quantities)
gg1
#gg1 + xlim(0,10) # limit x-axis from 0 to 10 generations
```

