---
title: "HostSwitch"
author: "Valeria Trivellone, Sabrina B.L. Araujo, Bernd Panassiti"
date: 'created: 15.04.2021, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for the HostSwitch package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}

header-includes:
  - \usepackage[document]{ragged2e}
  
bibliography: Literature.bib
biblio-style: "apalike"
link-citations: true

abstract: |
  \justifing
  
  The 'HostSwitch' package uses an individual-based model to simulate dispersion and host switch events by a consumer. As a consequence od the adaptation to a new resource,the consumer may expand its host range. 
  The HostSwitch R package provides functions to explore host switch events on a novel resource over the time. Based on the Individual-Based Model (IBM) published in Araujo et al. [-@Araujo_2015], the main function of this R package simulates the colonization dynamics of consumers over resources that are randomly offered. The HostSwitch package was implemented by including all functionalities in the original IBM model as well as numerous additions, such as comparison and evaluation of simulations with several customized parameter options and an in-depth description of the parameters in an ecological context.
  
  We aim to reach a broader audience by providing an interactive visualization of the model and linking the theoretical model to real world scenarios. The dataset ("data$parli") provided with  the present package provide a useful series of example with some parameters retrived from the literature.\n\n\n
  

---

<style type="text/css">
h1  {
  margin-top: 1em;
}
h1.title {
  font-size: 38px;
  color: DarkBlue;
  text-align: center;
}

h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
</style>




```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8.5, fig.height=5.5, fig.align='center', warning=FALSE, message=FALSE, cache = T)
```


# Theoratical background

The evolution of the biological associations may include several concatenated events of speciation affecting one or both species and it is driven by four main processes: cospeciation, host switch, failure to speciate and "missing the boat" (Page 2002). While the prevalent paradigm wants the cospeciation to be the main process driving the coevolution of most of the biological associations, recent evidence showed that host-switching, favored by the intrinsic capability of adaptation in the sub-optimal environment, may explain rapid novel associations (i.e. colonization of novel hosts at the ecological time scale) eventually followed by speciation (at the evolutionary time) as well the observed incongruences of the paired phylogenies (see Brooks et al. 2019 for a review).


# Motivation 

The simulation model of HostSwitch aims to measure the dynamics of the host switch (extent and frequency) in the population of an organism (hereafter Consumer) that interacts with current and potential hosts (hereafter Resource) over the generations. A host switch implies that a Consumer may colonize new Resource, which in turn imposes selection pressure that impacts the Consumers' survival. The host-switching relies on a mechanism of ecological readjustment or ecological fitting, i.e. the capability of the Consumer to use a similar Resource even if sub-optimal (Jansen 1985, Agosta 2008). The fundamental aspect of the HostSwitch simulation model is to track, summarize and compare the dispersion and successful host switch events in a new Resource by the populations of the Consumer.

The HostSwitch package offers simple functions to evaluate the role of the single parameters included in the model in defining the probability and the amplitude of the host switch. The extension in R code allows the users to choose among different consumer-resource scenarios which accommodate all the symbiotic (sensu lato) associations.

This package will be expanded with further scenarios of biological interactions. Please feel free to contact the authors to contribute by making suggestions, or sending bug fixes or extension proposals. 



# Using HostSwitch

## Installing, loading and citing the package

Install the recent release of HostSwitch package using GitHub

```{r, eval=FALSE}
devtools::install_github(repo = "berndpanassiti/HostSwitch",build_vignettes = TRUE)
```

Loading and citation

```{r}
library(HostSwitch)
citation("HostSwitch")
```


## Interactive plot using Shiny
```{r, eval=F}
shinyHostSwitch()
```

This function opens a new window with an interactive plot of the HostSwitch simulation model. This plot can be used to interact with the model and explore how different settings affect dispersion and host switch events by a consumer.

## Run the model
### Simulate host switches by consumers
```{r}
# using only 20 generations and 1 simulation, others arguments left to default values
# see ?simHostSwitch for more arguments
simulated_quantities = simHostSwitch(seed=123,n_sim=1,n_generation=20) 
```

The function simHostSwitch generates a list containing all quantities of interest which can be used for summary statistics or plots. Quantities of interest are optimum phenotypes that Consumers should have to be favored by the current Resource (**pRes_sim**), that Consumers should have to be favored by the novel Resource (**pRes_new_sim**), individual phenotype values of the Consumers (**pInd**), number of migrating individuals at each generation (**pInd_jump_sim**), individual phenotype values of the Consumers who disperse in a novel Resource (**pInd_whichjump_sim**), and individual phenotype values of the Consumers who successful colonize a novel Resource (**pInd_whichsurv_sim**). These parameters are available for each generation step.


### Summary of host switches simulation
```{r}
summaryHostSwitch(simulated_quantities) 
```

The function summaryHostSwitch creates a summary of basic statistics for phenotypes, dispersion and host switch events. 

### Plot the simulation of parasite host switches
```{r}
gg1 = plotHostSwitch(simulated_quantities, sim_n = 1) # use n_sim to select a specific simulation
gg1
#gg1 + xlim(0,10) # limit x-axis from 0 to 10 generations
```


## Comparing two host switch simulations

To investigate the effect of selected parameters on host switches, the testHostSwitch function can be used to compare two simulations.


Here, we investigate how different number of offsprings (parameter b) affect the average number of jumps to a new host by a parasite.  
First, we create two simulations with different number of offsprings, leaving other paramters to default values. Note, to test for differences between simulations it is necessary to choose reasonable high number of simulations. We recommand for the number of jumps to set n_sim >= 100. Second, we compare both simulations choosing the parameter of interest and wether or not a boxplot should be drawn.
```{r}
# see ?simHostSwitch for more arguments
m1 = simHostSwitch(seed=123,n_sim=100,b=10) 
m2 = simHostSwitch(seed=123,n_sim=100,b=15) 

# see ?testHostSwitch for more arguments
testHostSwitch(simulated_quantities1=m1,simulated_quantities2=m2,parameter="j",test="t",plot=FALSE)
```


# References
